---
title: "Report liver clinical"
author: "Ente Kang"
date: "`r Sys.Date()`"
output: word_document
---

# 3.1 Summary statistics of the major variables

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(survival)
library(reportRmd)
liv <- read_csv("K:/Xuan_Li/Ente_Kang/liver_clinic.csv")  #keep barcode for merging only, not in model
liv <- liv %>% select(-c(1))
rm_covsum(liv, covs = c("gender", "age", "stage"), maincov = "OS")
```
We see that for individuals who were censored, there was a smaller proportion of females than in the un-censored case. 

**Note: The above table and the analysis below were conducted using OS time, as there are more true survival outcomes (113). **

```{r, echo=FALSE, message=FALSE}
library(survminer)

f <- survfit(Surv(OS.time, OS) ~ gender, data = liv)
ggsurvplot(f, conf.int = T, pval = T, surv.median.line = "hv") +
  labs(y = "Overall Survival Time", x = "Time")

g <- survfit(Surv(OS.time, OS) ~ stage, data = liv)
ggsurvplot(g, conf.int = T, pval = T, surv.median.line = "hv") +
  labs(y = "Overall Survival Time", x = "Time")
```

Since there are 113 true survival outcomes, we should not entertain more than 11 variables. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
cn <- read_csv("K:/Xuan_Li/Ente_Kang/analysis plan/week 6/cn_genomic_pcs.csv")
dm <- read_csv("K:/Xuan_Li/Ente_Kang/analysis plan/week 6/dm_genomic_pcs.csv")
ge <- read_csv("K:/Xuan_Li/Ente_Kang/analysis plan/week 6/ge_genomic_pcs.csv")

allpcs <- cn %>% inner_join(dm, by="barcode") %>% inner_join(ge, by="barcode")

liv_final <- liv %>% inner_join(allpcs, by="barcode")

liv_final <- liv_final %>% mutate(stage_recode = ifelse(stage == "Stage I" | stage == "Stage II", "Early Stage", 
                                           "Not Early Stage"))

h <- survfit(Surv(OS.time, OS) ~ stage_recode, data = liv_final)
ggsurvplot(h, conf.int = T, pval = T, surv.median.line = "hv", legend.labs = c("Early Stage", "Not Early Stage")) +
  labs(y = "Overall Survival Probability", x = "Time (Days)")


```

```{r, echo=FALSE}
cox_small <- coxph(Surv(OS.time, OS) ~ age+gender+stage_recode, data = liv_final)
small_sum <- summary(cox_small)
sumdf <- data.frame(round(small_sum$coefficients, 3))
colnames(sumdf)[5] <- "p-value"
outTable(sumdf, caption = "Cox PH using vars. from clinical data only", row.names = "variables")
```

In the above model with age, gender and stage_recode, we see that stage_recode is highly significant. 
Suggests that stage of disease may be key. 



```{r, echo=FALSE}

rm_covsum(liv_final, covs = c("pc1_cn", "pc2_cn", "pc3_cn", "pc4_cn", "pc5_cn", "pc6_cn", 
                              "pc1_dm", "pc2_dm", "pc3_dm", "pc4_dm", "pc5_dm", "pc6_dm", 
                              "pc1_ge", "pc2_ge", "pc3_ge", "pc4_ge", "pc5_ge", "pc6_ge"), 
          maincov = "OS", caption = "Summary of first 6 pcs for 3 genomic datasets")
```



# Models

## Cox model CN

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(survival)
library(quantreg)
coxcn <- coxph(Surv(OS.time, OS) ~ gender+age+stage_recode+
                 pc1_cn+pc2_cn+pc3_cn+pc4_cn+pc5_cn+pc6_cn, data = liv_final)
cnsum <- summary(coxcn)
cnsum1 <- data.frame(round(cnsum$coefficients, 3))
colnames(cnsum1)[5] <- "p-value"
outTable(cnsum1, row.names = "variables", caption="Summary table for the CN cox model")
```

It appears that only stage_recode and pc1_cn were statistically significant.

## Cox model GE

```{r, echo=FALSE}
coxge <- coxph(Surv(OS.time, OS) ~ gender+age+stage_recode+
                 pc1_ge+pc2_ge+pc3_ge+pc4_ge+pc5_ge+pc6_ge, data = liv_final)
gesum <- summary(coxge)
gesum1 <- data.frame(round(gesum$coefficients, 3))
colnames(gesum1)[5] <- "p-value"
outTable(gesum1, row.names = "variables", caption="Summary table for the GE cox model")
```

In this univariate model ge cox ph, age, stage_recode were statistically significant.

## Cox model DM

```{r, echo=FALSE}
coxdm <- coxph(Surv(OS.time, OS) ~ gender+age+stage_recode+
                 pc1_dm+pc2_dm+pc3_dm+pc4_dm+pc5_dm+pc6_dm, data = liv_final)
dmsum <- summary(coxdm)
dmsum1 <- data.frame(round(dmsum$coefficients, 3))
colnames(dmsum1)[5] <- "p-value"
outTable(dmsum1, row.names = "variables", caption="Summary table for the DM cox model")
```

In the DM Cox PH model, only stage_recode was statistically significant. 

**Findings: stage_recode was key in all 3 models. The pcs didn't play much of a role, except for pc1_cn. We saw previously that the first 6 pcs were only able to explain ~50% of the variation**

## Cox model (full)

```{r, echo=FALSE, message=FALSE}

liv_final <- subset(liv_final, select = -c(barcode, stage, DSS, DSS.time, gender_n, agegroup, stagegroup))

cox1 <- coxph(Surv(OS.time, OS) ~., data = liv_final)
csum <- summary(cox1)
csum1 <- data.frame(round(csum$coefficients, 3))
colnames(csum1)[5] <- "p-value"
outTable(csum1, row.names = "variables", caption="Summary table for the full cox model")
```

## stepwise selection
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(MASS)
step <- stepAIC(cox1, direction = "both", trace = 0)
stepsum <- round(data.frame(summary(step)$coefficients),3)
colnames(stepsum)[5] <- "p-value"
outTable(stepsum, row.names = "variables", caption = "Stepwise selection results")
```

**Why is it that pc5_dm was chosen, and not pc3_dm or pc4_dm? Similarly, what about pc1_ge, pc2_ge**  
*It is to my understanding that the first few pcs are supposed to explain more variation*

## backward elimination
```{r, echo=FALSE, warning=FALSE, message=FALSE}
stepb <- stepAIC(cox1, direction = "backward", trace = 0)
bsum <- round(data.frame(summary(stepb)$coefficients),3)
colnames(bsum)[5] <- "p-value"
outTable(bsum, row.names='variables', caption='Backward elimination results')
```


### Cox PH assumption (reduced model)
```{r, echo=FALSE}
cox2 <- coxph(Surv(OS.time, OS) ~age+pc1_cn+pc1_dm+pc2_dm+pc5_dm+pc3_ge+pc4_ge+stage_recode, data = liv_final)
pha <- cox.zph(cox2)
d <- round(data.frame(pha$table),2)
outTable(d, row.names = 'variables', caption = 'Cox Proportional Hazards Assumption')
```

We see that there are no issues with the cox proportional hazards assumption


## Quantile Regression (only using clinical vars.) 
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(quantreg)
crq_small <- crq(Surv(OS.time, OS) ~age+gender+stage_recode, taus = c(0.25, 0.5),data = liv_final, method = "PengHuang")

tau_50s <- round(data.frame(summary(crq_small, tau=0.5)[[1]][2]), 3)
tau_25s <- round(data.frame(summary(crq_small, tau=0.25)[[1]][2]), 3)
colnames(tau_25s) <- c("Coefficient", "LB", "UB", "Se", "T stat", "p-value")
colnames(tau_50s) <- c("Coefficient", "LB", "UB", "Se", "T stat", "p-value")

outTable(tau_25s, row.names = "variables", caption="CQR (25th percentile), clinical vars.")
outTable(tau_50s, row.names = "variables", caption="CQR (50th percentile), clinical vars.")

```

**The coefficients seem less "wild" if we don't include the pcs. However, some values are still quite strange.**


## Quantile Regression (reduced, vars. from Cox)
```{r, echo=FALSE, warning=FALSE, message=FALSE}
crqmod <- crq(Surv(OS.time, OS) ~age+pc1_cn+pc1_dm+pc2_dm+pc5_dm+pc3_ge+pc4_ge+stage_recode, taus = c(0.25, 0.5),data = liv_final, method = "PengHuang")

tau_50 <- round(data.frame(summary(crqmod, tau=0.5)[[1]][2]), 3)
tau_25 <- round(data.frame(summary(crqmod, tau=0.25)[[1]][2]), 3)
colnames(tau_25) <- c("Coefficient", "LB", "UB", "Se", "T stat", "p-value")
colnames(tau_50) <- c("Coefficient", "LB", "UB", "Se", "T stat", "p-value")

outTable(tau_25, row.names = "variables", caption="Reduced CQR (25th percentile)")
outTable(tau_50, row.names = "variables", caption="Reduced CQR (50th percentile)")

```

**At the 25%-tile, only stage_recode and pc1_cn were statistically significant.**  

**At the 50%-tile, stage_recode and pc2_dm were significant. Why is pc2_dm significant but pc1_dm not? Isn't pc1_dm supposed to explain "more variation" in a sense?**

Do we need to consider rescaling age? The magnitude for age is much greater than that of the pcs. The coefs seem very large for CQR, but were okay for Coxph.  
Furthermore what are the units for survival time, is it days?